---
title: "Mulivariate Viz"
---

```{r}
library(tidyverse)

# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)

```

## Exercise 1
```{r}
# Exercise 1
ggplot(education, aes(x=sat)) +
  geom_histogram(color="white")

```
The SAT score averages in this plot range from ~840 to ~1110. The distribution of this data has a left skew.

## Exercise 2
```{r}
# Exercise 2
# Construct a plot of sat vs expend
# Include a "best fit linear regression model" (HINT: method = "lm")
ggplot(education, aes(y=sat, x=expend)) +
  geom_point() +
  geom_smooth(method = "lm")

# Construct a plot of sat vs salary
# Include a "best fit linear regression model" (HINT: method = "lm")
ggplot(education, aes(y=sat, x=salary)) +
  geom_point() +
  geom_smooth(method = "lm")

```
Surprisingly, there appears to be a negative correlation between a state's average SAT score and its average expenditure per student or average teacher salary.

## Exercise 3
```{r}
# Exercise 3
ggplot(education, aes(y=sat, x=salary, color=expend)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Exercise 4
```{r}
# Exercise 4
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 2))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```
As the average teacher salary and average expenditure per student in a state increases, the state's average SAT score decreases.

## Exercise 5
```{r}
# 5a: Univariate visualization of fracCat
ggplot(education, aes(x=fracCat)) +
  geom_bar(color="white") +
  labs(y = "Number of U.S. states")
```
```{r}
# 5b: 2 bivariate visualizations that demonstrate the relationship between sat and fracCat
ggplot(education, aes(y=sat, x=fracCat)) +
  geom_boxplot()

ggplot(education, aes(x=sat, fill=fracCat)) +
  geom_density(alpha = 0.5)
```
This graph tells me that there is a negative correlation between how many students take the SAT and a state's average SAT score.

```{r}
# 5c: Trivariate visualization that demonstrates the relationship of sat with expend AND fracCat
ggplot(education, aes(y = sat, x = expend, color = fracCat)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```
This graph makes clear spending does not have a major impact on a state's average SAT score. The primary determinant of a state's average SAT score is actually the fraction of eligible students who take the exam.

It appeared that SAT scores decreased as spending increased because I was excluding important information from my early visualizations of this data. Only by considering more variables was I able to understand the actual trends occurring in the data.
